{\bf Landslide: A New Race-Finding Tool for 15-410}

### 2018-09-27 irb/landslide-guide-f18.tex
copy lecture for f18; make it non-research; add non-research user guide

\section{Introduction}

You should have received a document titled {\em User Study Information Sheet} that describes your rights as a study
participant and our confidentiality procedures. If you are missing this, please contact me immediately. My email
address is {\tt bblum@cs.cmu.edu} and my phone number is 412-304-4294; email is preferred.
\\

\noindent Likewise, please make sure you have read and understand the lecture slides on systematic testing, which are
available on the 15-410 course website. If you have any questions about the research background I would be
happy to answer them.

@@

We encourage you to use Landslide on the GHC cluster or UNIX pool machines ({\tt ghcXX.ghc.andrew.cmu.edu}
or {\tt unix.andrew.cmu.edu}). Landslide has not been tested with other configurations and will probably fail
mysteriously on your personal laptop, and we cannot promise to be of any help.

\begin{enumerate}
	\item Clone the repository - {\tt git clone https://github.com/bblum/landslide.git} - this will be your workspace.
	\item Run {\tt ./p2-setup.sh PATH}, where {\tt PATH} is the absolute path to your p2 implementation.
		\begin{itemize}
			\item This command will import your p2 into {\tt ./pebsim/p2-basecode/} and then attempt to build it.
				Whenever you update your p2 code, please do so in its original {\tt PATH}
				and then run {\tt p2-setup.sh} again.
			\item {\tt PATH} should be the path with {\tt 410kern}, {\tt 410user}, {\tt user/libthread}, etc as subdirectories, not the {\tt user/libthread} directory itself.
		\end{itemize}
	\item Run {\tt ./landslide OPTIONS} to run tests with Landslide, for various values of {\tt OPTIONS}:
		\begin{enumerate}
			\item Use ./landslide -h to view command line options. The most important ones are -t , max testing
time, and -p , which test program to run.
			\item Landslide's bug reports will show up as HTML files.
				\begin{itemize}
					\item If you are using a cluster machine in person, you can view them by pointing a browser to \\
						{\tt file:///PATH/TO/landslide/FILENAME.html}
					\item For remote usage, move the html file to {\tt ~/www} in your AFS, and view it at \\
						{\tt http://www.contrib.andrew.cmu.edu/~YOUR\_ANDREWID/FILENAME.html}.
						Note the www in the URL is mandatory;
						note also that this will allow the world to see it,
						so be careful to clean up any traces in {\tt ~/www} when you're done!
				\end{itemize}
			\item Suggested test configurations to run are as follows.
				You can, of course, change the time limits or test programs as you see fit.
				For the longer-running configurations,
				we recommend running them inside of a {\tt screen} or {\tt tmux} session.
				\begin{itemize}
					\item \texttt{./landslide -t 30m -p thr\_exit\_join}
					\item \texttt{./landslide -t 30m -p mutex\_test}
					\item \texttt{./landslide -t 30m -p paradise\_lost}
					\item \texttt{./landslide -t 30m -p broadcast\_test}
					\item \texttt{./landslide -t 30m -p paraguay}
					\item \texttt{./landslide -t 30m -p rwlock\_downgrade\_read\_test}
						\\
					\item \texttt{./landslide -t 3h -p thr\_exit\_join}
					\item \texttt{./landslide -t 3h -p mutex\_test}
					\item \texttt{./landslide -t 3h -p paradise\_lost}
					\item \texttt{./landslide -t 3h -p broadcast\_test}
					\item \texttt{./landslide -t 3h -p paraguay}
					\item \texttt{./landslide -t 3h -p rwlock\_downgrade\_read\_test}
						\\
					\item \texttt{./landslide -t 12h -p thr\_exit\_join}
					\item \texttt{./landslide -t 12h -p mutex\_test}
					\item \texttt{./landslide -t 12h -p paradise\_lost}
					\item \texttt{./landslide -t 12h -p broadcast\_test}
					\item \texttt{./landslide -t 12h -p paraguay}
					\item \texttt{./landslide -t 12h -p rwlock\_downgrade\_read\_test}
				\end{itemize}
			\item To build your own custom test case for use with Landslide,
				put it in {\tt pebsim/p2-basecode/user/progs}
				and edit the {\tt STUDENTTESTS} line in {\tt pebsim/p2-basecode/config-incomplete.mk}
				(note: NOT {\tt config.mk}; the setup script auto-generates that file).
				Then run {\tt ./p2-setup.sh} again.
		\end{enumerate}
	\item After fixing any found bugs, re-run Landslide with the same test options to confirm that your fix works.
\end{enumerate}

@@

\noindent {\bf If something goes wrong:} If you find a Landslide bug, such as an assertion/crash, or such as a
bug report that you think is actually a bug in Landslide rather than a bug in your P2, or such as it getting stuck
and not exiting in the given time limit, please create a tarball of your workspace, just as you would to report
a reference kernel bug, but send it to Ben ({\tt bblum@cs.cmu.edu}) instead of 410 staff. For any other technical
issues getting it to work, or if you have difficulty understanding a bug report, send email to Ben.
\\

\noindent However, we cannot provide help on how to understand why your bug happened or how to fix it. The point of
the study is that Landslide's automation plus your own brain should be enough! On the other hand, if you have
a design question while deciding between potential ways to fix a bug, send it to 410 course staff as usual.

>>>>

\section{Preface}

%You should have received a document titled {\em User Study Information Sheet} that describes your rights as a study
%participant and our confidentiality procedures. If you are missing this, please contact me immediately. My email
%address is {\tt bblum@cs.cmu.edu} and my phone number is 412-304-4294; email is preferred.
%\\

Please make sure you have read and understand the lecture slides on systematic testing, which are
available on the 15-410 course website. If you have any questions about the research background I would be
happy to answer them; you can get in touch at {\tt bblum@cs.cmu.edu}.

@@

We encourage you to use Landslide on the GHC cluster or UNIX pool machines ({\tt ghcXX.ghc.andrew.cmu.edu}
or {\tt unix.andrew.cmu.edu}).
%Landslide has not been tested with other configurations and will probably fail
%mysteriously on your personal laptop, and we cannot promise to be of any help.
Since Landslide uses the open-source simulator Bochs rather than the proprietary Simics,
you can also try it on any personal machine which runs Linux.
(The Landslide repository includes the Bochs source; you do not need to install it separately.)

\begin{enumerate}
	\item Clone the repository - {\tt git clone https://github.com/bblum/landslide.git} - this will be your workspace.
	\item Run {\tt ./p2-setup.sh PATH}, where {\tt PATH} is the absolute path to your p2 implementation.
		\begin{itemize}
			\item This command will import your p2 into {\tt ./pebsim/p2-basecode/} and then attempt to build it.
				Whenever you update your p2 code, please do so in its original {\tt PATH}
				and then run {\tt p2-setup.sh} again.
			\item {\tt PATH} should be the path with {\tt 410kern}, {\tt 410user}, {\tt user/libthread}, etc as subdirectories, not the {\tt user/libthread} directory itself.
			\item This step may take a few minutes, as it needs to compile the Bochs simulator from source.
		\end{itemize}
	\item Run {\tt ./landslide OPTIONS} to run tests with Landslide, for various values of {\tt OPTIONS}:
		\begin{enumerate}
			\item Use ./landslide -h to view command line options. The most important ones are -t , max testing
time, and -p , which test program to run.
			\item Suggested test configurations to run are as follows.
				You can, of course, change the time limits or test programs as you see fit.
				\begin{itemize}
					\item \texttt{./landslide -t 30m -p thr\_exit\_join}
					\item \texttt{./landslide -t 30m -p mutex\_test}
					\item \texttt{./landslide -t 30m -p paradise\_lost}
					\item \texttt{./landslide -t 30m -p broadcast\_test}
					\item \texttt{./landslide -t 30m -p paraguay}
					\item \texttt{./landslide -t 30m -p rwlock\_downgrade\_read\_test}
						\\
					%\item \texttt{./landslide -t 3h -p thr\_exit\_join}
					%\item \texttt{./landslide -t 3h -p mutex\_test}
					%\item \texttt{./landslide -t 3h -p paradise\_lost}
					%\item \texttt{./landslide -t 3h -p broadcast\_test}
					%\item \texttt{./landslide -t 3h -p paraguay}
					%\item \texttt{./landslide -t 3h -p rwlock\_downgrade\_read\_test}
					%	\\
					\item \texttt{./landslide -t 12h -p thr\_exit\_join}
					\item \texttt{./landslide -t 12h -p mutex\_test}
					\item \texttt{./landslide -t 12h -p paradise\_lost}
					\item \texttt{./landslide -t 12h -p broadcast\_test}
					\item \texttt{./landslide -t 12h -p paraguay}
					\item \texttt{./landslide -t 12h -p rwlock\_downgrade\_read\_test}
				\end{itemize}
				Please check slide 34 of the lecture before attempting to run {\tt cyclone}, {\tt racer}, {\tt largetest}, etc.

				For the longer-running configurations,
				we recommend running them inside of a {\tt screen} or {\tt tmux} session.
				The GHC and UNIX cluster machines reboot nightly, so using a personal machine is recommended
				for running tests overnight.
			\item Landslide's bug reports will show up as HTML files.
				\begin{itemize}
					\item If you are using a cluster machine in person, or a personal machine,
						you can view them by pointing a browser to \\
						{\tt file:///PATH/TO/landslide/FILENAME.html}
					\item For remote usage, move the html file to {\tt ~/www} in your AFS, and view it at \\
						{\tt http://www.contrib.andrew.cmu.edu/~YOUR\_ANDREWID/FILENAME.html}.
						Note the www in the URL is mandatory;
						note also that this will allow the world to see it,
						so be careful to clean up any traces in {\tt ~/www} when you're done!
				\end{itemize}
			\item To build your own custom test case for use with Landslide,
				put it in {\tt pebsim/p2-basecode/user/progs}
				and edit the {\tt STUDENTTESTS} line in {\tt pebsim/p2-basecode/config-incomplete.mk}
				(note: NOT {\tt config.mk}; the setup script auto-generates that file).
				Then run {\tt ./p2-setup.sh} again.
		\end{enumerate}
	\item After fixing any found bugs, re-run Landslide with the same test options to confirm that your fix works.
\end{enumerate}

@@

\section{Troubleshooting}

If you find a Landslide bug, such as an assertion/crash, or such as a
bug report that you think is actually a bug in Landslide rather than a bug in your P2, or such as it getting stuck
and not exiting in the given time limit, please create a tarball of your workspace, just as you would to report
a reference kernel bug, but send it to Ben ({\tt bblum@cs.cmu.edu}) instead of 410 staff. For any other technical
issues getting it to work, or if you have difficulty understanding a bug report, send email to Ben.
\\

\noindent
This is the first semester we have released Landslide with Bochs instead of Simics,
so there may be any manner of unexpected bugs (it's fairly well tested on my own laptop, though!).
If you are trying Landslide on a personal machine and have problems with the setup step,
please try again on one of the official CMU servers before sending me a bug report.
If all else fails, you may try the known-good Simics version:
\url{https://github.com/bblum/landslide-simics}
(which requires a GHC or UNIX cluster machine for the software license).
\\

\noindent Finally, we cannot provide help on how to understand why your bug happened or how to fix it. The point of
the research is that Landslide's automation plus your own brain should be enough! On the other hand, if you have
a design question while deciding between potential ways to fix a bug, send it to 410 course staff as usual.

### 2018-09-27 landslide-lecture/2018f/lecture.tex
copy lecture for f18; make it non-research; add non-research user guide

\begin{document}
\renewcommand{\inserttotalframenumber}{39}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		%\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Big Picture: Execution Tree Exploration}
	\textbf{Backtracking}
	\begin{itemize}
		\item Each time test completes, identify a PP to replay differently
		\item Reset machine state and start over
		\item Implemented using Simics bookmarks
			\begin{itemize}
				\item {\tt set-bookmark} and {\tt skip-to}
			\end{itemize}
		\item Replay test from the beginning, with a different interleaving
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Test Suite}
	Landslide ships with 6 approved test cases:
	\linegap

	Standard P2 tests
	\begin{itemize}
		\item {\tt thr\_exit\_join}
		\item {\tt paraguay}
		\item {\tt rwlock\_downgrade\_read\_test}
	\end{itemize}

	New tests
	\begin{itemize}
		\item {\tt broadcast\_test}
		\item {\tt paradise\_lost} %(new this semester)
		\item {\tt mutex\_test} %(new this semester)
	\end{itemize}
\end{frame}

@@

\begin{frame}{User Study}
	\textbf{Try Landslide on your P2!}
	\begin{itemize}
		\item Bare minimum effort: No more than 1 hour
		\begin{itemize}
			\item Clone a github URL, run setup script, run tests, answer survey
			\item Landslide will automatically report test results (as described below)
		\end{itemize}
		\item Full study plan: 4-8 hours of active attention
		\begin{itemize}
			\item (Estimated, including time to diagnose and fix bugs)
			\item However, many tests should run passively overnight -- start soon!
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{User Study - Additional Information}
	Human Subjects Research
	\begin{itemize}
		\item CMU IRB has approved this study
		\item Landslide will collect results while you use it
		\begin{itemize}
			\item Record commands issued, take snapshots of your P2 code
			\item All data will be anonymized before publication
		\end{itemize}
		\item No coercion: {\bf \em There is no penalty for not participating.}
		\begin{itemize}
			\item I am not on course staff, cannot influence your grade
			\item Course staff will not have access to study data during semester
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Risks \& Benefits
	\begin{itemize}
		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
		\item Risk: Landslide may find no bugs and be a waste of your time.
		\item Benefit: You might learn something...
	\end{itemize}
	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
\end{frame}

\begin{frame}{User Study - How to Participate}
	Interested?
	\linegap

	To participate\dots
	\begin{itemize}
		%\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item (watch your email for the link)
				%\item \url{http://tinyurl.com/landslide-p2-s17}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Next week, room and time TBD
			\end{itemize}
		%\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	%% im out of time :<<
	%\linegap

	%{\em Really} interested?
	%\begin{itemize}
	%	\item I know of an M.S. thesis topic if you want to do post-OS research.
	%\end{itemize}
\end{frame}

>>>>

\begin{document}
\renewcommand{\inserttotalframenumber}{38}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	{\bf About Landslide}
	\begin{itemize}
		%\item Simics module, which traces:
		\item Simulator* module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
				\item *Used to be Simics, now Bochs (open-source, 3x faster)
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		%\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Big Picture: Execution Tree Exploration}
	\textbf{Backtracking}
	\begin{itemize}
		\item Each time test completes, identify a PP to replay differently
		\item Reset machine state and start over
		% ehh.. don't feel like explaining
		%\item Implemented using {\tt fork()}
		%\item Implemented using Simics bookmarks
		%	\begin{itemize}
		%		\item {\tt set-bookmark} and {\tt skip-to}
		%	\end{itemize}
		\item Replay test from the beginning, with a different interleaving
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Test Suite}
	6 of the official P2 tests are ``Landslide-friendly'':
	\begin{itemize}
		\item {\tt thr\_exit\_join}
		\item {\tt paraguay}
		\item {\tt rwlock\_downgrade\_read\_test}
		\item {\tt broadcast\_test}
		\item {\tt paradise\_lost} %(new this semester)
		\item {\tt mutex\_test} %(new this semester)
	\end{itemize}
\end{frame}

@@

\begin{frame}{Try Landslide on your P2!}
	Bare minimum effort: No more than 1 hour
	\begin{itemize}
		\item Clone a github URL, run setup script, run tests, answer survey
		\item Landslide will automatically report test results (as described below)
	\end{itemize}
	\linegap

	Full study plan: 4-8 hours of active attention
	\begin{itemize}
		\item (Estimated, including time to diagnose and fix bugs)
		\item However, many tests should run passively overnight -- start soon!
	\end{itemize}
	\pause
	\linegap

@@

%\begin{frame}{User Study - Additional Information}
%	\sout{Human Subjects Research} not anymore!
%	\begin{itemize}
%		\item CMU IRB has approved this study
%		\item Landslide will collect results while you use it
%		\begin{itemize}
%			\item Record commands issued, take snapshots of your P2 code
%			\item All data will be anonymized before publication
%		\end{itemize}
%		\item No coercion: {\bf \em There is no penalty for not participating.}
%		\begin{itemize}
%			\item I am not on course staff, cannot influence your grade
%			\item Course staff will not have access to study data during semester
%		\end{itemize}
%	\end{itemize}
%	\pause
%	\linegap
%
%	Risks \& Benefits
%	\begin{itemize}
%		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
%		\item Risk: Landslide may find no bugs and be a waste of your time.
%		\item Benefit: You might learn something...
%	\end{itemize}
%	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
%\end{frame}

\begin{frame}{Try Landslide on your P2!}
	Interested?
	\linegap

	To participate\dots
	\begin{itemize}
		%\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Read \& follow user guide instructions
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/landslide-guide-f18.pdf}
				\item (this will also be emailed to you)
				%\item \url{http://tinyurl.com/landslide-p2-s17}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Next week, room and time TBD
			\end{itemize}
		%\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
\end{frame}
